A hybrid approach for news recommender system using optimization methods

### 2021-03-01 paper/paper.tex
added content based filtering

    Common approaches, such as collaborative filtering, has its own problems: cold start, scalability and data sparsity. Content-based approaches suffer from the fact that we have to somehow represent recommended item in feature space.

@@

\begin{itemize}
    \item \textbf{Collaborative filtering}: generating predictions about the interests of a user by collecting preferences or taste information from other users. It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person
    \item \textbf{Content-based filtering}:
    \item \textbf{Session filtering}:
    \item \textbf{Popularity filtering}:
    \item \textbf{Demographic filtering}:
    \item \textbf{Time-based filtering}:
\end{itemize}

@@

Candidate generation stage not only filters items, but also attaches weight to items.

@@

@@

    \begin{table}[h]
        \centering
        \begin{tabular}{ccc}
            \toprule
            \textit{user\_id} & \textit{content\_based\_filtering} & \textit{collaborative\_filtering} \\
            \midrule
            1                 & 0.25                                  & 1                               \\
            2                 & 1                               & 0.5                                \\
            \vdots & \vdots & \vdots \\
            \bottomrule
            \end{tabular}%
        \caption{boost values}
        \label{tab:boost_values}
    \end{table}

>>>>

    Common approaches, such as collaborative filtering, have their own problems: cold start, scalability and data sparsity. Content-based approaches suffer from the fact that we have to somehow represent recommended item in feature space.

@@

\begin{itemize}
    \item \textbf{Collaborative filtering}: generating predictions about the interests of a user by collecting preferences or taste information from other users. It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person
    \item \textbf{Content-based filtering}: is based on the assumption that people who liked items with certain attributes in the past, will like the same kind of items in the future as well. It makes use of item features to compare the item with user profiles and provide recommendations
    \item \textbf{Session filtering}: is selecting cadidates based on user's activity on current session
    \item \textbf{Popularity filtering}: uses information such as number of views, shows, comments, etc.
    \item \textbf{Demographic filtering}: uses demographic data such as age, gender, education, etc. for identifying categories of users
    \item \textbf{Time-based filtering}: ranking news is a way that more recent items have higher scores
\end{itemize}

@@

Candidate generation stage not only filters items, but also attaches weights to them.

@@

We are representing each news item as a vector of confidence values (how strong each item belongs to category).

\begin{table}[h]
    \centering
    \begin{tabular}{cccccc}
        \toprule
        item\_id & \textit{politics} & \textit{IT} & \textit{social\_media} & $\cdots$ & \textit{confilcts} \\
        \midrule
        1 & 0.55 & 0 & 0 & $\cdots$ & 0.3 \\
        2 & 0    & 0.81  & 0.62 & $\cdots$ & 0 \\
        \vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\

        \bottomrule
    \end{tabular}%
    \caption{text vectors}
    \label{tab:text_vectors}
\end{table}

Tagging is made in the following way: we have a dictionary of words that belongs to categories:

\begin{table}[h]
    \centering
    \begin{tabular}{cc}
        \toprule
        \textit{word} & \textit{category} \\
        \midrule
        alexey\_nalalny & politics \\
        crash & accident \\
        telegeram & IT \\
        \vdots & \vdots \\
        \bottomrule
    \end{tabular}%
    \caption{words' categories}
    \label{tab:words_categories}
\end{table}

Confidence is taken from word's tf-idf metric alongside all corpus.

As user interacts with the system, we are forming his preferences vector in the following way:

\begin{equation}
    \texttt{user\_vector} \mathrel{+}= \texttt{action\_score} * \texttt{item\_vector}
\end{equation}

Where action score is taken from table:

\begin{table}[h]
    \centering
    \begin{tabular}{cc}
        \toprule
        \textit{action} & \textit{score} \\
        \midrule
        shown but now viewed & -1 \\
        viewed & 0.5 \\
        emoji or comment & 0.5 \\
        read till the end & 1 \\
        \bottomrule
    \end{tabular}%
    \caption{actions' scores}
    \label{tab:action_score}
\end{table}

As we now have both user's and items' vectors, we are able to find similarities between them:

\begin{equation}
    r_i = 1 - \cos{(\texttt{user\_vector}, \texttt{item\_vector})}
\end{equation}

@@

    \begin{table}[]
        \centering
        \begin{tabular}{ccc}
            \toprule
            \textit{user\_id} & \textit{content\_based\_filtering} & \textit{collaborative\_filtering} \\
            \midrule
            1                 & 0.25                                  & 1                               \\
            2                 & 1                               & 0.5                                \\
            \vdots & \vdots & \vdots \\
            \bottomrule
            \end{tabular}%
        \caption{boost values}
        \label{tab:boost_values}
    \end{table}
